Social Visual Analytics: DevOps for Data Scientists

### 2015-03-25 introduction.tex
intro work

Over fifty years ago, John Tukey published ``The Future of Data Analysis''
in which he defined a remarkably farsighted agenda for the future of
data analysis and statistical computing. His vision has been realized
through environments like S and R, that started a revolution by
integrating statistical graphics, computation, and robust techniques
for practical data modeling. Every day, many thousands of data scientists
and others specialists benefit from this astounding vision.

The field has now moved beyond the boundaries of what Tukey originally
foresaw. Processing and networking capabilities today far exceed
what was barely imagined fifty years ago. Inexpensive, scalable
virtualized systems allow analysts to routinely solve problems
involving millions or even billions of data values. Networking
also enables both collaboration and the automation of complex
tasks and processes in ways that were not considered even ten years ago.
We turn our attention to this opportunity to use computation
to support, not only individuals, but entire teams and their
work within larger organizations.

As a starting point, consider the role of a data science team within
a business or technical organization today.
Analysts work in loose or tight collaboration with colleagues and
domain experts.
Project teams vary in size, from just a few people to dozens or
more, even within one project's duration. Assignments are often
broad, and include tasks such as problem identification,
data wrangling, modeling, analysis, visualization, summarization,
presentation and interpretation of results, and recommending
actions to help clients to realize the benefits of the analysis.
Eventually, knowledge or working prototypes created by data science teams
may be transferred to other organizations to employ them in production.
There are many details to these tasks. For example, data wrangling can involve
finding data, understanding its syntax and semantics, assessing data quality,
performing normalization and data quality remediation, and making the data
available to other tasks that will follow.

>>>>

Almost a
half-century ago, Tukey foresaw
much of what is now commonplace in data analysis~\cite{TukeyFDA}.
Powerful, interactive environments for analysis and programming were
the goal, together with an unflinching (and, at the time, somewhat
heretical) insistence in keeping humans as a central part of the
discovery process. His now-famous quip that ``the picture-examining eye is the
best finder we have of the wholly unanticipated'' has come to define
much of visual analytics and exploratory visualization~\cite{TukeyEDA}.

In some way, we have moved far beyond what Tukey imagined:
computing and networking capabilities today far exceed
what was barely imaginable then. 
We argue, on the other hand, that we \emph{develop} our data analysis
solutions, and context in which they become part of the broader issues
being solved in organizations, has not changed as much: the S language
was developed essentially alongside Unix~\cite{TheSSystem}, after all.
In this paper, we turn our attention to this opportunity to use computation
to support, not only individuals, but entire teams and their
work within larger organizations.

Consider the role of a data science team within
a business or technical organization today.
Project teams vary in size, from just a few people to dozens or
more, even within one project's duration. Assignments are often
broad, and include tasks such as problem identification,
data wrangling, modeling, analysis, visualization, summarization,
presentation and interpretation of results, and recommending
actions to help clients to realize the benefits of the analysis.
Eventually, knowledge or working prototypes created by these teams
are transferred to other organizations to employ them in production.
There are many details to these tasks. For example, data wrangling can involve
finding data, understanding its syntax and semantics, assessing data quality,
performing normalization and data quality remediation, and making the data
available to other tasks that will follow.
